Active Learning for Level Set Estimation

### 2013-05-01 ijcai13pres.tex
Finish changes discussed with Andreas

@@

\begin{frame}
\begin{center}
\uncover<1->{Swimmers of Lake Zurich, beware!}
\vspace{0.2in}
\begin{columns}[c]
\column{0.4\textwidth}
\uncover<2->{
\sig{\includegraphics[width=1.7in]{figures/pr.jpeg}}{www.limnobotics.ch}\\[1em]
\sig{\includegraphics[width=1.7in]{figures/bloom.jpg}}{Flickr/Dr. Jennifer L. Graham/U.S. Geological Survey}
}
\column{0.6\textwidth}
\minipage[c][0.80\textheight][s]{\columnwidth}
\uncover<1->{
\small
``The warming waters of one of central Europe's most popular holiday
destinations, Switzerland's Lake Zurich, have created an ideal environment
for a population explosion of algae including \emph{Planktothrix rubescens},
[\ldots]''
\qauth{Scientific American}
}
\vfill
\uncover<3->{
\small
``\emph{Planktotrhix rubescens} can account for half of the total
phytoplankton biomass in Lake Zurich in summer [\ldots]''
\vspace{1em}

``\emph{Planktotrhix rubescens} are among the most important
producers of hepatotoxic microcystins in freshwaters [\ldots]''
\qauth{Silke Van den Wyngaert \emph{et al.}, ASLO, 2011}
}
\vfill
\uncover<4->{
\small
``Microcystins [\ldots] are cyanotoxins and can be very toxic for plants and
animals including humans.
Their hepatotoxicity may cause serious damage to the liver.''
\qauth{Wikipedia}
}
\endminipage
\end{columns}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{0.2in}
\only<1>{
\sig{\includegraphics[width=2in]{figures/manual.jpg}}{www.limnobotics.ch}
}
\only<2>{
\sig{\includegraphics[width=4in]{figures/boat.jpg}}{www.limnobotics.ch}
}
\only<3>{
\sig{\includegraphics[width=4in]{figures/lake.png}}{www.limnobotics.ch}
}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{0.2in}
\includegraphics<1>[width=4.7in]{figures/limno_bgape_sc}
\includegraphics<2>[width=4.7in]{figures/limno_bgape}
\includegraphics<3>[width=4.7in]{figures/limno_bgape_ls}
\only<4>{\hspace{-2em}\includegraphics[width=4.45in]{figures/limno_bgape_cl}}
\end{center}
\end{frame}

\begin{frame}
Pose as a sequential decision making problem (\emph{pool-based active learning}):
\begin{itemize}
\item<1-> No measurements available in advance, just a set (pool) of possible\\sampling locations ($D$)
\vspace{0.5em}
\item<2->[]\hspace{-1.5em} At each iteration $t \geq 1$:
\item<2-> Decide where to measure next ($\*x_t \in D$)
\item<3-> Obtain noisy observation ($y_t = f(\*x_t) + n_t$)
\item<4-> Update our estimate of $f$
\end{itemize}
%\begin{center}
%\color{white}
%\includegraphics<1>[draft,width=4.45in]{figures/oned_0}
%\color{black}
%\includegraphics<2>[width=4.45in]{figures/oned_0}
%\includegraphics<3>[width=4.45in]{figures/oned_1_0}
%\includegraphics<4>[width=4.45in]{figures/oned_1_1}
%\includegraphics<5>[width=4.45in]{figures/oned_1_2}
%\includegraphics<6>[width=4.45in]{figures/oned_2_0}
%\includegraphics<7>[width=4.45in]{figures/oned_2_1}
%\includegraphics<8>[width=4.45in]{figures/oned_2_2}
%\end{center}
\end{frame}

\begin{frame}
\begin{enumerate}
\item<1-> How do we \textbf{estimate} the function?
\vspace{1em}
\item<2-> How do we \textbf{classify}?
\vspace{1em}
\item<3-> Each measurement is expensive (time, battery power).\\
      How do we \textbf{select} ``informative'' measurements?
\end{enumerate}
\vspace{2em}
\begin{center}
\uncover<4->{\large
Gaussian processes to the rescue!
}
\end{center}
\end{frame}

@@

\begin{frame}
\begin{itemize}
\item<1-> What if we don't know the threshold level $h$?
\vspace{1em}
\item<2-> Implicitly define $h = \omega \max_{\*x\in D} f(\*x),\ 0 < \omega < 1$
\vspace{1em}
\item<3-> No existing algorithms for this problem (to our knowledge)
\vspace{1em}
\item<4-> We extend \acl (and its theory!) to this setting
\vspace{1em}
\item<5->\iacl algorithm:
\begin{itemize}
\item<6-> Need to accurately estimate the maximum $\rightarrow$ modified selection rule
\item<7-> $h$ is now an estimated quantity $\rightarrow$ modified classification rules
\end{itemize}
\end{itemize}
\begin{center}
%\color{white}
%\only<1-2>{\includegraphics[draft,width=3.77in]{figures/voned_cl_41_imp_1}}
%\color{black}
%\only<3>{\includegraphics[width=3.77in]{figures/voned_cl_41_imp_1}}
%\only<4->{\includegraphics[width=3.77in]{figures/voned_cl_41_imp_2}}
\end{center}
\end{frame}

@@

\begin{frame}
\begin{itemize}
\item<1-> Up to this point we have assumed a fixed cost per sample
\item<2-> What about the traveling distance between measurements?
\item<3-> We extend \acl to select a \emph{batch} of sampling locations at each step (\bacl)
\item<3-> Use this to plan ahead:
\begin{itemize}
\item Select a batch of sampling locations
\item Connect them using a Euclidean TSP path
\item Traverse path and collect measurements
\end{itemize}
\end{itemize}

\begin{center}
\color{white}
\includegraphics<1-3>[draft,height=1.8in]{figures/limno_bgape_pp_0_0}
\color{black}
\includegraphics<4>[height=1.8in]{figures/limno_bgape_pp_0_0}
\includegraphics<5>[height=1.8in]{figures/limno_bgape_pp_0_1}
\includegraphics<6>[height=1.8in]{figures/limno_bgape_pp_0_2}
\includegraphics<7>[height=1.8in]{figures/limno_bgape_pp_30_0}
\includegraphics<8>[height=1.8in]{figures/limno_bgape_pp_30_1}
\includegraphics<9>[height=1.8in]{figures/limno_bgape_pp_30_2}
\includegraphics<10>[height=1.8in]{figures/limno_bgape_pp_60_0}
\includegraphics<11>[height=1.8in]{figures/limno_bgape_pp_60_1}
\includegraphics<12>[height=1.8in]{figures/limno_bgape_pp_60_2}
\includegraphics<13>[height=1.8in]{figures/limno_bgape_pp_90_0}
\includegraphics<14>[height=1.8in]{figures/limno_bgape_pp_90_1}
\includegraphics<15>[height=1.8in]{figures/limno_bgape_pp_90_2}
\includegraphics<16>[height=1.8in]{figures/limno_bgape_pp_120_0}
\includegraphics<17>[height=1.8in]{figures/limno_bgape_pp_120_1}
\includegraphics<18>[height=1.8in]{figures/limno_bgape_pp_120_2}
\includegraphics<19>[height=1.8in]{figures/limno_bgape_pp_150_0}
%\includegraphics<20>[height=1.8in]{figures/limno_bgape_pp_150_1}
%\includegraphics<21>[height=1.8in]{figures/limno_bgape_pp_150_2}
%\includegraphics<22>[height=1.8in]{figures/limno_bgape_pp_180_0}
%\includegraphics<23>[height=1.8in]{figures/limno_bgape_pp_180_1}
%\includegraphics<24>[height=1.8in]{figures/limno_bgape_pp_180_2}
%\includegraphics<25>[height=1.8in]{figures/limno_bgape_pp_210_0}
\includegraphics<20>[height=1.8in]{figures/ev_bgape_pp}
\end{center}
\end{frame}

@@

\section*{Conclusion}

\begin{frame}
\begin{center}
\large Summary
\end{center}
\vspace{-1em}
\begin{itemize}
\item<1-> \acl algorithm:
\begin{itemize}
\item<2-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Theoretical guarantees & \includegraphics[width=1.3in]{figures/thm.png} \end{tabular}
\item<3-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Competitive with the state of the art in practice (sometimes considerably superior) & \hspace{-0.5em}\includegraphics[width=1.3in]{figures/ev_bgape_seq_long} \end{tabular}
\end{itemize}
\item<4-> Two useful extensions:
\begin{itemize}
\item<5-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Implicit threshold level & \includegraphics[width=1.3in]{figures/limno_bgape_imp_class_140} \end{tabular}
\item<6-> \begin{tabular}{m{5.7cm} >{\centering}m{4cm}}Batch sampling & \includegraphics[width=1.3in]{figures/limno_bgape_pp_30_2} \end{tabular}
\end{itemize}
\item<7-> Look out for algae when swimming in Lake Zurich! {\color{blue}$\ddot\smile$}
\end{itemize}
\end{frame}

>>>>

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

@@

\begin{frame}
\begin{center}
\uncover<1->{Swimmers of Lake Zurich, beware!}
\vspace{0.2in}
\begin{columns}[c]
\column{0.4\textwidth}
\uncover<2->{
\sig{\includegraphics[width=1.7in]{figures/pr.jpeg}}{www.limnobotics.ch}\\[1em]
\sig{\includegraphics[width=1.7in]{figures/bloom.jpg}}{Flickr/Dr. Jennifer L. Graham/U.S. Geological Survey}
}
\column{0.6\textwidth}
\minipage[c][0.80\textheight][s]{\columnwidth}
\uncover<1->{
\small
``[\ldots]Switzerland's Lake Zurich [\ldots] an ideal environment
for a population explosion of algae including \emph{Planktothrix rubescens}
[\ldots]''
\qauth{Scientific American}
}
\vfill
\uncover<3->{
\small
``\emph{Planktotrhix rubescens} are among the most important
producers of hepatotoxic microcystins in freshwaters [\ldots]''
\qauth{Silke Van den Wyngaert \emph{et al.}, ASLO, 2011}
}
\vfill
\uncover<4->{
\small
``Microcystins [\ldots] are cyanotoxins and can be very toxic for plants and
animals including humans.
Their hepatotoxicity may cause serious damage to the liver.''
\qauth{Wikipedia}
}
\endminipage
\end{columns}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{0.2in}
%\only<1>{
%\sig{\includegraphics[width=2in]{figures/manual.jpg}}{www.limnobotics.ch}
%}
\only<1>{
Autonomous surface vehicle developed by the Autonomous Systems Lab of ETH\\[1em]
\sig{\includegraphics[width=4in]{figures/boat.jpg}}{www.limnobotics.ch}
}
\only<2>{
Take measurements on a vertical transect of the lake\\[1em]
\sig{\includegraphics[width=4in]{figures/lake.png}}{www.limnobotics.ch}
}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{0.2in}
\only<1>{
Original algae concentration measurements ($\sim 2000$)\\[1em]
\includegraphics[width=4.7in]{figures/limno_bgape_sc}
}
\only<2>{
Interpolated algae concentration field\\[1em]
\includegraphics[width=4.7in]{figures/limno_bgape}
}
\only<3>{
Focus on accurately estimating regions of ``high'' concentration (e.g. $\geq 7$)\\[1em]
\includegraphics<3>[width=4.7in]{figures/limno_bgape_ls}
}
\only<4>{
Classify transect into a {\color{red!40!yellow}super-} and a {\color{locol}sub}level set\\[1em]
\hspace{-2em}\includegraphics[width=4.45in]{figures/limno_bgape_cl}
}
\end{center}
\end{frame}

\begin{frame}
Pose as a sequential decision making problem (\emph{pool-based active learning}):
\begin{itemize}
\item<2-> No measurements available in advance, just a set (pool) of possible\\sampling locations ($D$)
\vspace{0.5em}
\item<3->[]\hspace{-1.5em} At each iteration $t \geq 1$:
\item<3-> Decide where to measure next ($\*x_t \in D$)
\item<4-> Obtain noisy observation ($y_t = f(\*x_t) + n_t$)
\item<5-> Update our estimated classification
\end{itemize}
%\begin{center}
%\color{white}
%\includegraphics<1>[draft,width=4.45in]{figures/oned_0}
%\color{black}
%\includegraphics<2>[width=4.45in]{figures/oned_0}
%\includegraphics<3>[width=4.45in]{figures/oned_1_0}
%\includegraphics<4>[width=4.45in]{figures/oned_1_1}
%\includegraphics<5>[width=4.45in]{figures/oned_1_2}
%\includegraphics<6>[width=4.45in]{figures/oned_2_0}
%\includegraphics<7>[width=4.45in]{figures/oned_2_1}
%\includegraphics<8>[width=4.45in]{figures/oned_2_2}
%\end{center}
\end{frame}

\begin{frame}
\begin{enumerate}
\item<1-> How do we \textbf{estimate} the underlying function from measurements?
\vspace{1em}
\item<2-> How do we \textbf{classify}?
\vspace{1em}
\item<3-> Each measurement is expensive (time, battery power).\\
      How do we \textbf{select} ``informative'' measurements?
\end{enumerate}
\vspace{2em}
\begin{center}
\uncover<4->{\large
Gaussian processes to the rescue!
}
\end{center}
\end{frame}

@@

\begin{frame}
\begin{center}
\large Implicit threshold level
\end{center}
\begin{itemize}
\item<1-> What if we don't know the threshold level $h$?
\vspace{1em}
\item<2-> Implicitly define $h = \omega \max_{\*x\in D} f(\*x),\ 0 < \omega < 1$
\vspace{1em}
\item<3-> No existing algorithms for this problem (to our knowledge)
\vspace{1em}
\item<4-> We extend \acl (and its theory!) to this setting
\vspace{1em}
\item<5->\textbf{\iacl} algorithm:
\begin{itemize}
\item<6-> Need to accurately estimate the maximum $\rightarrow$ modified selection rule
\item<7-> $h$ is now an estimated quantity $\rightarrow$ modified classification rules
\end{itemize}
\end{itemize}
\begin{center}
%\color{white}
%\only<1-2>{\includegraphics[draft,width=3.77in]{figures/voned_cl_41_imp_1}}
%\color{black}
%\only<3>{\includegraphics[width=3.77in]{figures/voned_cl_41_imp_1}}
%\only<4->{\includegraphics[width=3.77in]{figures/voned_cl_41_imp_2}}
\end{center}
\end{frame}

@@

\begin{frame}
\begin{center}
\large Batch sampling
\end{center}
\begin{itemize}
\item<1-> Up to this point we have assumed a fixed cost per sample.\\
          What about the traveling distance between measurements?
\vspace{1em}
\item<2-> We extend \acl to select a \emph{batch} of sampling locations at each step
\vspace{1em}
\item<3-> Plan ahead:
\begin{itemize}
\item Use \textbf{\bacl} to select a batch of sampling locations
\item Connect them using a Euclidean TSP path
\item Traverse path and collect measurements
\end{itemize}
\end{itemize}

\begin{center}
\color{white}
\includegraphics<1-3>[draft,height=1.8in]{figures/limno_bgape_pp_0_0}
\color{black}
\includegraphics<4>[height=1.8in]{figures/limno_bgape_pp_0_0}
\includegraphics<5>[height=1.8in]{figures/limno_bgape_pp_0_1}
\includegraphics<6>[height=1.8in]{figures/limno_bgape_pp_0_2}
\includegraphics<7>[height=1.8in]{figures/limno_bgape_pp_30_0}
\includegraphics<8>[height=1.8in]{figures/limno_bgape_pp_30_1}
\includegraphics<9>[height=1.8in]{figures/limno_bgape_pp_30_2}
\includegraphics<10>[height=1.8in]{figures/limno_bgape_pp_60_0}
\includegraphics<11>[height=1.8in]{figures/limno_bgape_pp_60_1}
\includegraphics<12>[height=1.8in]{figures/limno_bgape_pp_60_2}
\includegraphics<13>[height=1.8in]{figures/limno_bgape_pp_90_0}
%\includegraphics<14>[height=1.8in]{figures/limno_bgape_pp_90_1}
%\includegraphics<15>[height=1.8in]{figures/limno_bgape_pp_90_2}
%\includegraphics<16>[height=1.8in]{figures/limno_bgape_pp_120_0}
%\includegraphics<17>[height=1.8in]{figures/limno_bgape_pp_120_1}
%\includegraphics<18>[height=1.8in]{figures/limno_bgape_pp_120_2}
%\includegraphics<19>[height=1.8in]{figures/limno_bgape_pp_150_0}
%\includegraphics<20>[height=1.8in]{figures/limno_bgape_pp_150_1}
%\includegraphics<21>[height=1.8in]{figures/limno_bgape_pp_150_2}
%\includegraphics<22>[height=1.8in]{figures/limno_bgape_pp_180_0}
%\includegraphics<23>[height=1.8in]{figures/limno_bgape_pp_180_1}
%\includegraphics<24>[height=1.8in]{figures/limno_bgape_pp_180_2}
%\includegraphics<25>[height=1.8in]{figures/limno_bgape_pp_210_0}
\includegraphics<14>[height=1.8in]{figures/ev_bgape_pp}
\end{center}
\end{frame}

\section{Conclusion}

@@

\begin{frame}
\begin{center}
\large Summary
\end{center}
\vspace{-1.5em}
\begin{itemize}
\item<2-> \acl algorithm:\vspace{0.5em}
\begin{tabular}{>{\centering\arraybackslash} m{5cm}>{\centering\arraybackslash} m{5.5cm}}
Theoretical guarantees & Competitive with the state of the art\\
\includegraphics[height=0.9in]{figures/thm.png} & \includegraphics[height=1in]{figures/ev_chl_seq_long}
\end{tabular}
\item<3-> Two useful extensions:\vspace{0.5em}
\begin{tabular}{>{\centering\arraybackslash} m{5.25cm}>{\centering\arraybackslash} m{5.25cm}}
Implicit threshold level (\iacl) & Batch sampling (\bacl)\\
\includegraphics[height=0.95in]{figures/limno_bgape_imp_class_140} & \includegraphics[height=0.95in]{figures/limno_bgape_pp_30_2}
\end{tabular}
\item<4-> Look out for algae when swimming in Lake Zurich! {\color{blue}$\ddot\smile$}
\end{itemize}
\end{frame}
